A Methodology to Define QoS and SLA Requirements in Service Choreographies

### 2012-08-09 articles/CAMAD2102/article-version2.tex
More fixing and improving

 This paper presents a new approach to detect non-functional failures of the participants in a choreography during the design stage. Our approach
 conducts a performance evaluation by means of generalized stochastic Petri nets (GSPN) as the intermediate representation of a choreography
 specified in BPMN2, a notational language to model business process, with the representation of the QoS model. Our work differs from 
 those found in the literature because it is employed in stages before the development (for example, in the modeling and design). With our 
 proposal it is possible to establish requirements for QoS and SLA in early stages in order to plan the capacity of the network connecting
 the hosts involved in the deployment of the application, reducing the costs and unexpected behaviors  in later stages of the development.

@@

  The two main approaches to compose services are choreographies and orchestrations. A service choreography allow the composition of services
 in a collaborative manner. A choreography describes the P2P interactions of the externally observable behavior of its participants. Different
 from orchestrations, choreographies don't have a single point of control or coordination \cite{Barker2009}. Choreography tend to be more
 efficient than orchestrations because the probability to have a bottleneck in the system is reduced \cite{Barker2009b}.

@@

Recently, \cite{Xia2011} proposed an approach to predict the reliability of composite services specified in WS-CDL. It is also
 presented a complete translation of a choreography specified in WS-CDL to a GSPN. Similar to \cite{Xia2011}, we also use the formalism
 of Petri nets as an intermediate representation of a choreography. \cite{Xia2011} is focused in reliability, while our proposal is focused in
 QoS attributes of communication network and we use the interaction model of BPMN2 as choreography language.  
%differ from \cite{Xia2011} in terms of the objective. We define QoS requirements, while

@@

The contribution of this paper is to present a new methodology for evaluate the performance of choreographies with the objective of
 define  QoS requirements. The evaluation is performed in the design stage of the choreography by using GSPNs. The proposal is based 
on the execution of the following steps:

@@

\end{enumerate}

@@

    \begin{mydef}
      A choreography of BPMN processes is a tuple: \\
       $PC = (\mathcal{O, A, E, G, T, S}, \{e^S\}, \mathcal{E}^I, \{e^E\}, \mathcal{E}^{I_M}, \mathcal{E}^{I_T}, \mathcal{G}^F, \mathcal{G}^J,$
       $\mathcal{G}^X, \mathcal{G}^M, \mathcal{G}^V, \mathcal{F} )$ where:
      \begin{itemize}
	\item $\mathcal{O}$ is the set of objects, which is partitioned in three disjoint sets: activities ($\mathcal{A}$), events $\mathcal{E}$
	and \emph{gateways} $\mathcal{G}$.

	\item $\mathcal{A}$, is the set of choreography tasks and it is the same as $\mathcal{T}$.

@@

	\item $\mathcal{G}$ é the set of gateways and it is partitioned in the 	five disjoints sets of parallel fork gateways $\mathcal{G}^F$,
	parallel join gateways $\mathcal{G}^J$, data-based XOR gateways $\mathcal{G}^X$, XOR merge gateways $\mathcal{G}^V$ and event-based
	XOR gateways $\mathcal{G}^M$.
      \end{itemize}
    \end{mydef}

@@

	\FOR { $CT_i \in \mathcal{T} $ }
	  \STATE $GSPN_{QoS} \leftarrow GSPN_{QoS} \oplus PNQoS(CT_i)$
          \STATE Add a arrival timed Transition at first of the $GSPN_{QoS}$.
	\ENDFOR

@@

	\STATE Add a starting Place and immediate Transition at the beginning of the $GSPN_{QoS}$.
        \STATE Add a ending Place and immediate Transition the end of the $GSPN_{QoS}$.

@@

%Nesta seção são aplicados os passos descritos da nossa abordagem no cenário de exemplo da Seção \ref{serviceChoreography} .
To attest the efficacy of our proposal, in this section all the steps of our methodology are performed. The choreography
 presented in Figure~\ref{fig:choreographyExample} is used in the process  because that scenario is simple in order to show and understand better the
 proposed methodology, and includes at least one element of three choreography  BPMN types (task, gateways and events). Each one of the
 next subsections describes one of the three steps of  the proposal defined in the Section~\ref{sec:proposta}.

@@

\begin{table}[!h]
    \centering
    %\caption{Descrição de transições da GSPN do exemplo }
    \caption{  Description of the transitions in the GSPN of Figure~\ref{fig:ExampleChoreographyGSPNQoS} }
    \label{table:transition_description}
    \begin{tabular}{|l|l|}
      \hline
       Transition     &       Description \\
       \hline
       $T_{Starting}$	 &  Starting choreography instance. \\
       $T_{Ending}$      &  Ending choreography instance. \\
       $T_{send1}$, $T_{send2}$ e $T_{send3}$   &  Activity of message sending. \\
       $T_{arrival1}$, $_2$, $_3$  &  Choreography instances arrival.\\
       $T_{arrival2}$, $_3$   &  Flow decision.\\
       $T_{latency1}$, $_2$, $_3$   &     Network latency\\
       $T_{receive1}$, $_2$, $_3$   &     Receive and service execution.\\	
       $T_{cerror1}$, $_2$, $_3$  &  Communication error. \\
       $T_{merror1}$, $_2$, $_3$ &  Format fault message. \\
       $T_{g}$   & Flow sequence continuation. \\
       $T_{joining2}$, $_3$   &   Flow joining. \\

@@

The next step is to define the weights representing the timed transitions and the priorities in the immediate transitions. These values
 depend on the knowledge about the environment where the choreography will be executed and the reliability of the independent
 services. In this paper we don't consider uncertainties in these values. Table~\ref{table:transitionsConfigurations} presents
 the definition of the weights in two scenarios, that will be needed to the simulations on the next step. Scenario 1 is configured
 to have less probability of failures than Scenario 2.

\begin{table}[!h]
    \centering
    %\caption{Configuração de pesos nos Cenários 1 e 2}
    \caption{Weights of Scenario 1 and Scenario 2}
    \label{table:transitionsConfigurations}
    \begin{tabular}{|l|l|l|}
      \cline{1-3}
		    &   \multicolumn{2}{|c|}{ Weights } \\
      \cline{1-3}
      %\hline
       \multicolumn{1}{|c|}{Transition }     		&      Scenario 1   &   Scenario 2 		\\
       \hline
       $T_{latency1}$, $T_{latency2}$, $T_{latency3}$ 	& 	0.99       &   0.94	\\
       $T_{cerror1}$, $T_{cerror2}$, $T_{cerror3}$      	&  	0.01 	   &   0.06	\\
       $T_{receive1}$, $T_{receive2}$, $T_{receive3}$  &  	99  	   &    97	\\
       $T_{merr1}$, $T_{merr2}$, $T_{merr3}$   &   1 	   &    3	\\
       $T_{arrival2}$, $T_{arrival3}$   &  0.5  &	0.5	\\
      \hline
    \end{tabular}
\end{table}

@@

To simulate the scenarios, 100 tokens are considered to each scenario at the place Start. Each token represent one instance of execution
 of the simulation. So, in each simulation, 100 concurrent instances of the scenarios were executed. Since there are probabilities
 involved in the experiments, as seen in Table~\ref{table:transitionsConfigurations}, the experiments need to be executed several times. There were defined 1500
 fires and 10 replications. To simulate the scenarios, the simulator Pipe2 \cite{Bonet2007} was used.

 Table~\ref{table:simulationsAll} shows the results of the simulations. The numbers presented in the table show the average of the tokens 
that arrived at each one of the places of the GSPN.

>>>>

 This paper presents a new approach to detect non-functional failures of the participants in a choreography during the design stage. Our approach
 conducts a performance evaluation by means of generalized stochastic Petri nets (GSPNs) as the intermediate representation of a choreography
 specified in BPMN2, a notational language to model business process, with the representation of the QoS model. Our work differs from 
 those found in the literature because it is employed in stages before the development (for example, in the modeling and design). With our 
 proposal it is possible to establish requirements for QoS and SLA in early stages in order to plan the capacity of the network connecting
 the hosts involved in the deployment of the application, reducing the costs and unexpected behaviors  in later stages of the development.

@@

  The two main approaches to compose services are choreographies and orchestrations. A service choreography allow the composition of services
 in a collaborative manner. A choreography describes the P2P interactions of the externally observable behavior of its participants. Different
 from orchestrations, choreographies don't have a single point of control or coordination \cite{Barker2009}. Choreographies tend to be more
 efficient than orchestrations because the probability to have a bottleneck in the system is reduced \cite{Barker2009b}.

@@

In \cite{Xia2011} it is  proposed an approach to predict the reliability of composite services specified in WS-CDL. It is also
 presented a complete translation of a choreography specified in WS-CDL to a GSPN. Similar to \cite{Xia2011}, we also use the formalism
 of Petri nets as an intermediate representation of a choreography. Different from \cite{Xia2011}, our proposal is focused in 
 QoS attributes of communication network and we use the interaction model of BPMN2 as choreography language.  
%differ from \cite{Xia2011} in terms of the objective. We define QoS requirements, while

@@

The contribution of this paper is  a new methodology to evaluate the performance of choreographies with the objective of
 define  QoS requirements. The evaluation is performed in the design stage of the choreography by using GSPNs. The proposal is based 
on the execution of the following steps:

@@

%The simulations are employed to retrieve approximated performance measures with certain confidence level
\end{enumerate}

@@

    \begin{mydef}
      A choreography of BPMN processes is a tuple: \\
       $PC = (\mathcal{O, A, E, G, T}, \{e^S\}, \mathcal{E}^I, \{e^E\}, \mathcal{E}^{I_M}, \mathcal{E}^{I_T}, \mathcal{G}^F, \mathcal{G}^J,$
       $\mathcal{G}^X, \mathcal{G}^M, \mathcal{G}^V, \mathcal{F} )$ where:
      \begin{itemize}
	\item $\mathcal{O}$ is the set of objects, which is partitioned in three disjoint sets: activities ($\mathcal{A}$), events $\mathcal{E}$
	and \emph{gateways} $\mathcal{G}$.

	\item $\mathcal{A}$, is the set of \textit{Choreography Tasks} ($\mathcal{T}$), \textit{Sub-choreographies} and \textit{Call Choreographies}. However, 
	\textit{Sub-choreographies} and \textit{Call Choreographies} aren't taken into account in this work, then $\mathcal{A}$ it is the same as $\mathcal{T}$.

@@

	\item $\mathcal{G}$ é the set of gateways and it is partitioned in the 	five disjoints sets of parallel fork gateways $\mathcal{G}^F$,
	parallel join gateways $\mathcal{G}^J$, data-based XOR gateways $\mathcal{G}^X$, XOR merge gateways $\mathcal{G}^V$ and event-based
	XOR gateways $\mathcal{G}^M$.
	\item $\mathcal{F} \subseteq \mathcal{O}x\mathcal{O}$ is the control flow relation, i.e. a set of sequence flows connecting objects,
      \end{itemize}
    \end{mydef}

@@

	\FOR { $CT_i \in \mathcal{T} $ }
	  \STATE $GSPN_{QoS} \leftarrow GSPN_{QoS} \oplus PNQoS(CT_i)$
          \STATE Add a arrival timed Transition at beginning of the $GSPN_{QoS}$.
	\ENDFOR

@@

	\STATE Add a starting Place and immediate Transition at the beginning of the $GSPN_{QoS}$.
        \STATE Add a ending Place and immediate Transition at the end of the $GSPN_{QoS}$.

@@

%Nesta seção são aplicados os passos descritos da nossa abordagem no cenário de exemplo da Seção \ref{serviceChoreography} .
To attest the efficacy of our proposal, in this section all the steps of our methodology are performed. The choreography
 presented in Figure~\ref{fig:choreographyExample} is used in the process  because that scenario is simple in order to show and understand better the
 proposed methodology, and includes at least one element of the three choreography  BPMN types considered (activities, gateways and events). Each one of the
 next subsections describes one of the three steps of  the proposal defined in the Section~\ref{sec:proposta}.

@@

\begin{table}[!h]
    \centering
    %\caption{Descrição de transições da GSPN do exemplo }
    \caption{  Description of the transitions in the GSPN of Figure~\ref{fig:ExampleChoreographyGSPNQoS} }
    \label{table:transition_description}
    \begin{tabular}{|l|l|}
      \hline
       Transition     &       Description \\
       \hline
       $T_{arrival1}$  &  Choreography instances arrival.\\
       $T_{arrival2}$, $_3$   & Choreography instances arrival and  Flow decisions.\\
       $T_{send1}$, $T_{send2}$ e $T_{send3}$   &  Activity of message sending. \\
       $T_{latency1}$, $_2$, $_3$   &     Network latency\\
       $T_{receive}$, $_2$, $_3$   &     Receive and service execution.\\	
       $T_{cerr1}$, $_2$, $_3$  &  Communication error. \\
       $T_{merr1}$, $_2$, $_3$ &  Format fault message. \\
       $T_{g}$   & Flow sequence continuation. \\
       $T_{ok1}, T_{ok2}, T_{ok3}$   &  Valid receive message . \\
       $T_{joining2}$, $_3$   &   Flow joining. \\
       %$T_{starting}$	 &  Starting choreography instance. \\
       $T_{ending}$      &  Ending choreography instance. \\

@@

The next step is to define the weights (rates) representing the timed transitions and the priorities in the immediate transitions. These values
 depend on the knowledge about the environment where the choreography will be executed and the reliability of the independent
 services. In this paper we don't consider uncertainties in these values. Table~\ref{table:transitionsConfigurations} presents
 the definition of the weights and priorities in two scenarios, that will be needed to the simulations on the next step. Scenario 1 is configured
 to have less probability of failures than Scenario 2.

\begin{table}[!h]
    \centering
    %\caption{Configuração de pesos nos Cenários 1 e 2}
    \caption{Weights of Scenario 1 and Scenario 2}
    \label{table:transitionsConfigurations}
    \begin{tabular}{|l|l|l|}
      \cline{1-3}
		    &   \multicolumn{2}{|c|}{ Weights } \\
      \cline{1-3}
      %\hline
       \multicolumn{1}{|c|}{Transition }     		&      Scenario 1   &   Scenario 2 		\\
       \hline
       $T_{latency1}$, $T_{latency2}$, $T_{latency3}$ 	& 	0.99       &   0.94	\\
       $T_{cerr1}$, $T_{cerr2}$, $T_{cerr3}$      	&  	0.01 	   &   0.06	\\
       $T_{receive}$, $T_{receive2}$, $T_{receive3}$  &  	99  	   &    97	\\
       $T_{merr1}$, $T_{merr2}$, $T_{merr3}$   &   1 	   &    3	\\
       $T_{arrival2}$, $T_{arrival3}$   &  0.5  &	0.5	\\
      \hline
    \end{tabular}
\end{table}

@@

To simulate the scenarios, 100 tokens are considered to each scenario at the place Start. Each token represents one instance of execution
 of the simulation. So, in each simulation, 100 concurrent instances of the scenarios were executed (multiple-server semantic). 
%Since  there are probabilities  involved in the experiments
The two scenarios were executed several times according to configurations in Table~\ref{table:transitionsConfigurations}. There were defined 1500
 fires and 10 replications. The Pipe2 \cite{Bonet2007} tool was used to model GSPNs and simulate the scenarios.

 Table~\ref{table:simulationsAll} shows the results of the simulations with confidence level of $95\%$. The numbers presented in 
the table show the average of the tokens that arrived at each one of the places of the GSPN.
